Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-02-15 Codes/PrepareResults.rmd

```{r}
variogram_tb <-
  data.frame(
    Parameters = c("alpha_ij", "beta_ij", "ymax_ij", "varepsilon_ij"),
    Range = c(400, 400, 400, 400),
    Mean = c(-0.5, format(0), 12000, format(0)),
    Nugget = format(c(0, 0, 0, 0)),
    Sill = c(0.02, format(1, nsmall = 1), 2000000, 0.015)
  ) %>%
  flextable() %>%
  compose(i = 1, j = 1, part = "body", value = as_paragraph(as_i("\U03B1"), as_sub(as_i("i,j")))) %>%
  compose(i = 2, j = 1, part = "body", value = as_paragraph(as_i("\U03B2"), as_sub(as_i("i,j")))) %>%
  compose(i = 3, j = 1, part = "body", value = as_paragraph(as_i("ymax"), as_sub(as_i("i,j")))) %>%
  compose(i = 4, j = 1, part = "body", value = as_paragraph(as_i("\U03B5"), as_sub(as_i("i,j")))) %>%
  compose(i = 1, j = 5, part = "body", value = as_paragraph("2.0", "\U2A2F", "10", as_sup("-2"))) %>%
  compose(i = 3, j = 3, part = "body", value = as_paragraph("1.2", "\U2A2F", "10", as_sup("4"))) %>%
  compose(i = 3, j = 5, part = "body", value = as_paragraph("2.0", "\U2A2F", "10", as_sup("6"))) %>%
  compose(i = 4, j = 5, part = "body", value = as_paragraph("1.5", "\U2A2F", "10", as_sup("-2"))) %>%
  align(align = "center", part = "all") %>%
  align(j = 1, align = "left", part = "all") %>%
  # --- change headers labels--- #
  set_header_labels(values = list(
    Parameters = "Parameters",
    Range = "Range (m)",
    Mean = "Mean",
    Nugget = "Nugget",
    Sill = "Sill"
  )) %>%
  #- change the borders just for consistency with other tables -#
  hline_bottom(part = "all") %>%
  hline_top(part = "header") %>%
  footnote(
    value = as_paragraph("NOTE: Only in ", as_i("ymax"), " are consistent units of measurement used for Mean, Nugget, and Sill ", "(", as_i("kg"),")."),
    ref_symbols = NA
  ) %>%
  autofit()
```

@@

plot <-
  ggplot() +
  geom_sf(data = field_plot_sf) +
  geom_sf(data = ex_plot, fill = "green", size = 1) +
  coord_sf(expand = FALSE) +
  # theme_void() +
  ggtitle("12 \u00D7 12 subplots") +
  theme_figure

@@

```{r, dependson = "setup"}
field_Ndesign <-
  ggplot() +
  geom_sf(
    data = field_plot_sf,
    aes(fill = factor(rate)), size = 0,
    inherit.aes = FALSE,
  ) +
  scale_fill_viridis_d() +
  labs(fill = "Nitrogen rate\n  (kg/ha)") +
  ggtitle("(a) Trial Design") +
  theme_figure
```

@@

```{r, dependson = "setup"}
# === subplot-level === #
vis_yield_subplot <-
  ggplot(field_subplot_sf) +
  geom_sf(aes(fill = yield), size = 0) +
  scale_fill_viridis_c() +
  labs(fill = "Yield Level\n  (kg/ha)") +
  ggtitle("(b) Simulated Yield Level") +
  theme_figure
```

@@

```{r}
fig_y_optN <-
  res_y_test %>%
  .[Method %in% c("RF", "BRF", "CNN")] %>%
  ggplot(aes(x = rmse_y, y = rmse_optN)) +
  geom_point(size = 0.5) +
  facet_grid(Model ~ Method) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 0.7) +
  stat_regline_equation(
    label.x = 2050, label.y = 90,
    aes(label = ..rr.label..),
    size = 3
  ) +
  guides(
    fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype = guide_legend(keywidth = 3, keyheight = 1),
    colour = guide_legend(keywidth = 3, keyheight = 1)
  ) +
  ylim(NA, 95) +
  labs(y = " RMSE of EONR Estimation (kg/ha)") +
  labs(x = " RMSE of Yield Prediction (kg/ha)") +
  theme_few() +
  theme(
    strip.text.x = element_text(size = 12, face = "bold"),
    strip.text.y = element_text(size = 12, face = "bold"),
    legend.title = element_blank(),
    legend.text = element_text(size = 12, face = "bold"),
    legend.position = "bottom"
  ) +
  theme_figure

# print(fig_y_optN, preview = "docx")
# dfabline <- data.frame(x = 0:3500, y=0:3500)
```

@@

```{r, dependson = "source-results"}
#/*----------------------------------*/
#' ## Distribution of RMSE of EONR estimates
#/*----------------------------------*/
plot_dis_optN <-
  allML_summary_bySim %>%
  .[type=="test",] %>%
  # .[Method %in% c("RF", "BRF", "CF_base"), ] %>%
  .[Method == "CF_base", Method := "CF_base"] %>%
  .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
  )] %>%
  .[, Model := factor(Model,
    levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt")
  )] %>%
  ggplot() +
  geom_density(aes(x = rmse_optN, fill = Method), alpha = 0.7) +
  facet_wrap(~Model, ncol = 1) +
  labs(x = "RMSE (kg/ha)") +
  theme_dist

@@

## Profit loss summary and distributions
```{r}
piLoss_density <-
  allML_summary_bySim %>%
  .[type == "test", ] %>%
  .[Method == "CF_base", Method := "CF-base"] %>%
  .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
  )] %>%
  .[, Model := factor(Model,
    levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt")
  )] %>%
  ggplot() +
  geom_density(aes(x = pi_loss, fill = Method), alpha = 0.7) +
  labs(x = expression(paste(hat(pi)["def"], " ($/ha)"))) +
  facet_wrap(~Model, ncol = 1) +
  theme_dist
```

>>>>

```{r}
variogram_tb <-
  data.frame(
    Parameters = c("alpha_ij", "beta_ij", "ymax_ij", "varepsilon_ij"),
    Range = c(400, 400, 400, 400),
    Mean = c(-0.5, format(0), 12000, format(0)),
    Nugget = format(c(0, 0, 0, 0)),
    Sill = c(0.02, format(1, nsmall = 1), 2000000, 0.015)
  ) %>%
  flextable() %>%
  compose(i = 1, j = 1, part = "body", value = as_paragraph(as_i("\U03B1"), as_sub(as_i("i,j")))) %>%
  compose(i = 2, j = 1, part = "body", value = as_paragraph(as_i("\U03B2"), as_sub(as_i("i,j")))) %>%
  compose(i = 3, j = 1, part = "body", value = as_paragraph(as_i("ymax"), as_sub(as_i("i,j")))) %>%
  compose(i = 4, j = 1, part = "body", value = as_paragraph(as_i("\U03B5"), as_sub(as_i("i,j")))) %>%
  compose(i = 1, j = 5, part = "body", value = as_paragraph("2.0", "\U2A2F", "10", as_sup("-2"))) %>%
  compose(i = 3, j = 3, part = "body", value = as_paragraph("1.2", "\U2A2F", "10", as_sup("4"))) %>%
  compose(i = 3, j = 5, part = "body", value = as_paragraph("2.0", "\U2A2F", "10", as_sup("6"))) %>%
  compose(i = 4, j = 5, part = "body", value = as_paragraph("1.5", "\U2A2F", "10", as_sup("-2"))) %>%
  align(align = "center", part = "all") %>%
  align(j = 1, align = "left", part = "all") %>%
  # --- change headers labels--- #
  set_header_labels(values = list(
    Parameters = "Parameters",
    Range = "Range (m)",
    Mean = "Mean",
    Nugget = "Nugget",
    Sill = "Sill"
  )) %>%
  #- change the borders just for consistency with other tables -#
  hline_bottom(part = "all") %>%
  hline_top(part = "header") %>%
  footnote(
    value = as_paragraph("NOTE: Only in ", as_i("ymax"), " are consistent units of measurement used for Mean, Nugget, and Sill ", "(", as_i("kg"),")."),
    ref_symbols = NA
  ) %>%
  # set_caption(
  #   caption = "Table C.1: The parameters for the Variogram models",
  #   style = "Table Caption"
  # ) %>%
  autofit()
```

@@

plot <-
  ggplot() +
  geom_sf(data = field_plot_sf) +
  geom_sf(data = ex_plot, fill = "green", size = 1) +
  coord_sf(expand = FALSE) +
  # theme_void() +
  ggtitle("12 \u00D7 32 plots") +
  theme_figure

@@

```{r, dependson = "setup"}
field_Ndesign <-
  ggplot() +
  geom_sf(
    data = field_plot_sf,
    aes(fill = factor(rate)), size = 0,
    inherit.aes = FALSE,
  ) +
  scale_fill_viridis_d() +
  labs(fill = "Nitrogen rate\n  (kg/ha)") +
  ggtitle("(1) Trial Design") +
  theme_figure
```

@@

```{r, dependson = "setup"}
# === subplot-level === #
vis_yield_subplot <-
  ggplot(field_subplot_sf) +
  geom_sf(aes(fill = yield), size = 0) +
  scale_fill_viridis_c() +
  labs(fill = "Yield Level\n  (kg/ha)") +
  ggtitle("(2) Simulated Yield Level") +
  theme_figure
```

@@

```{r}
fig_y_optN <-
  res_y_test %>%
  .[Method %in% c("RF", "BRF", "CNN")] %>%
  ggplot(aes(x = rmse_y, y = rmse_optN)) +
  geom_point(size = 0.5) +
  facet_grid(Model ~ Method) +
  geom_smooth(method = "lm", se = FALSE, color = "red", size = 0.7) +
  stat_regline_equation(
    label.x = 2050, label.y = 90,
    aes(label = ..rr.label..),
    size = 3
  ) +
  guides(
    fill = guide_legend(keywidth = 1, keyheight = 1),
    linetype = guide_legend(keywidth = 3, keyheight = 1),
    colour = guide_legend(keywidth = 3, keyheight = 1)
  ) +
  ylim(NA, 95) +
  labs(y = " RMSE of EONR Estimation (kg/ha)") +
  labs(x = " RMSE of Yield Prediction (kg/ha)") +
  theme_dist
```

@@

```{r, dependson = "source-results"}
#/*----------------------------------*/
#' ## Distribution of RMSE of EONR estimates
#/*----------------------------------*/
plot_dis_optN <-
  allML_summary_bySim %>%
  .[type=="test",] %>%
  # .[Method %in% c("RF", "BRF", "CF_base"), ] %>%
  .[Method == "CF_base", Method := "CF-base"] %>%
  .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
  )] %>%
  .[, Model := factor(Model,
    levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt")
  )] %>%
  ggplot() +
  geom_density(aes(x = rmse_optN, fill = Method), alpha = 0.7) +
  facet_wrap(~Model, ncol = 1) +
  labs(x = "RMSE (kg/ha)") +
  theme_dist

@@

## Profit loss summary and distributions
```{r}
piLoss_density <-
  allML_summary_bySim %>%
  .[type == "test", ] %>%
  .[Method == "CF_base", Method := "CF-base"] %>%
  .[, Model := case_when(
    Model == "aby" ~ "Scenario: aby",
    Model == "abytt" ~ "Scenario: abytt",
    Model == "aabbyy" ~ "Scenario: aabbyy",
    Model == "aabbyytt" ~ "Scenario: aabbyytt"
  )] %>%
  .[, Model := factor(Model,
    levels = c("Scenario: aby", "Scenario: abytt", "Scenario: aabbyy", "Scenario: aabbyytt")
  )] %>%
  ggplot() +
  geom_density(aes(x = pi_loss, fill = Method), alpha = 0.7) +
  labs(x = expression(paste(hat(pi)[italic("def")], " ($/ha)"))) +
  facet_wrap(~Model, ncol = 1) +
  theme_dist
```

### 2022-02-15 Writing/appendix.rmd

# Appendix A: Causal Tree Example {#ctree-ex}

Figure \@ref(fig:ct-ex) shows an example of CT where the treatment is an increase in seed rate from $27,000$ seeds/ha (control) to $31,000$ seeds/ha (treated) with apparent electrical conductivity ($\textit{ecs}$)($dS/m$) and topographical $\textit{slope}$ ($degree$) as variables to explain the heterogeneity of the impact of the treatment impact.

```{r ct-ex, fig.width=5, fig.height=5, fig.align="center", fig.cap = "An illustrative example of causal tree"}
fancyRpartPlot(opfit, sub = "")
```

@@

# Appendix B: Understanding yield response functions better

Another way to writing Eq. \@ref(eq:eqn6) that may help understanding the yield response (Eq. \@ref(eq:eqn6)) better conceptually is as follows:

\begin{alinged}
f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) \cdot N)) + \varepsilon. 
\end{alinged}

@@

\begin{equation}
f(N_{i}, \mathbf{c}_{i}) = ymax(\mathbf{c}_{i})(1-exp(\alpha(\mathbf{c}_{i}) + \beta(\mathbf{c}_{i})\cdot N_{i})) + \varepsilon_{i}. 
\end{equation}

The simplified notation of this equation, where $c_i$ is dropped, is \@ref(eq:eqn6) in section <span style = "color: blue;"> section number here </span>.

<span style = "color: blue;"> Shun, number the equations </span>

# Appendix C: Parameter generation

## Base parameters $\alpha$, $\beta$, and $ymax$ {#par-gen}
Parameters $\alpha_{i,j}$, $\beta_{i,j}$, $ymax_{i,j}$, and $\varepsilon_{i,j}$ were modeled as spatially autocorrelated using unconditional Gaussian geostatistical simulation based on the spherical variogram model. The `gstat` package (version 2.0.6 [@gstat1; @gstat2]) in R was used for this purpose

Table \@ref(tab:variogram-parameters) shows the main variogram parameter values used in generating $\alpha_{i,j}$, $\beta_{i,j}$, $ymax_{i,j}$, and $\varepsilon_{i,j}$, which were chosen to generate yields consistent with those observed in experiments conducted by the DIFM project. (Codes used to generate parameter values are publicly accessible as a Mendeley dataset [@Kakimoto2022codes].) The range and nugget of all parameters were assumed to be $400$ m and 0. The means of $ymax$, $\alpha$, $\beta$, and $\varepsilon$ were assumed to be $12,000$ kg/ha, $-0.5$, $0$ and $0$, and their sills were assumed to be $2.0 \times 10^{6}$ kg/ha, $0.02$, $1.0$ and $0.015$. See Figure \@ref(fig:fig-parameters) for an example of the generated parameters.

<span style = "color: blue;"> Shun, combnine all the figures </span>

```{r variogram-parameters, tab.cap = "The Parameters for the Variogram Models"}
variogram_tb
```

```{r fig-parameters, fig.width=7, fig.height=9, fig.cap = "An Illustrative Example of Spatial Distributions of Field Characteristics", cache = TRUE}

field_alpha + field_beta + field_ymax + field_m_error + field_optN + plot_layout(ncol = 2)
```

## Splitting the base parameters {#split-pars}

@@

>>>>

<!-- NOTE: Erase the extra Figure, Table and section numbers manually-->

## Appendix A: Causal Tree Example {-}

Figure A.1 shows an example of CT where the treatment is an increase in seed rate from $27,000$ seeds/ha (control) to $31,000$ seeds/ha (treated) with apparent electrical conductivity ($\textit{ecs}$)($dS/m$) and topographical $\textit{slope}$ ($degree$) as variables to explain the heterogeneity of the impact of the treatment impact.

```{r ct-ex, fig.width=5, fig.height=5, fig.align="center", fig.cap = "Figure A.1: An illustrative example of a causal tree"}
par(family="Times")
fancyRpartPlot(opfit, sub="")
```

@@

## Appendix B: Understanding Yield Response Functions Better {-}

Another way to writing Eq. \@ref(eq:eqn3) that may help understanding the yield response Eq. (B.1) better conceptually is as follows:

$$
f(N, \mathbf{c}) = ymax(\mathbf{c})(1-exp(\alpha(\mathbf{c}) + \beta(\mathbf{c}) \cdot N)) + \varepsilon. \qquad \text{(B.1)}
$$

@@

$$
f(N_{i}, \mathbf{c}_{i}) = ymax(\mathbf{c}_{i})(1-exp(\alpha(\mathbf{c}_{i}) + \beta(\mathbf{c}_{i})\cdot N_{i})) + \varepsilon_{i}. \qquad \text{(B.2)}
$$

The simplified notation of this equation, where $c_i$ is dropped, is Eq. \@ref(eq:eqn3) in section 3.1.2.

## Appendix C: Parameter Generation {-}

### C.1 $\quad$ Base parameters **$\alpha$**, **$\beta$**, and **$ymax$** {-}
Parameters $\alpha_{i,j}$, $\beta_{i,j}$, $ymax_{i,j}$, and $\varepsilon_{i,j}$ were modeled as spatially autocorrelated using unconditional Gaussian geostatistical simulation based on the spherical variogram model. The `gstat` package (version 2.0.6 [@gstat1; @gstat2]) in R was used for this purpose

Table C.1.1 shows the main variogram parameter values used in generating $\alpha_{i,j}$, $\beta_{i,j}$, $ymax_{i,j}$, and $\varepsilon_{i,j}$, which were chosen to generate yields consistent with those observed in experiments conducted by the DIFM project. (Codes used to generate parameter values are publicly accessible as a Mendeley dataset [@Kakimoto2022codes].) The range and nugget of all parameters were assumed to be $400$ m and 0. The means of $\alpha$, $\beta$, $ymax$, and $\varepsilon$ were assumed to be $-0.5$, $0$, $12,000$ kg/ha, and $0$, and their sills were assumed to be $0.02$, $1.0$, $2.0 \times 10^{6}$ kg/ha, and $0.015$. See Figure C.1.1 for an example of the generated parameters.

```{r, tab.cap = "Table C.1.1: The parameters for the variogram models"}
variogram_tb
```

```{r, fig.width=7, fig.height=9, fig.cap = "Figure C.1.1: An illustrative example of spatial distributions of field characteristics", cache = TRUE}
field_alpha + field_beta + field_ymax + field_m_error + field_optN + plot_layout(ncol = 2)
```

### C.2 $\quad$ Splitting the Base Parameters {-}

@@

## Appendix D:  Training Accuracy of the ML methods {-}

During the simulations, we evaluated the training accuracy using the same ML models we presented in the Result and Discussion section. 
Table D.1 shows the mean RMSE of the EONR estimation and $\hat{\pi}_{def}$ over the one thousand simulations using the training datasets as testing datasets. Overall, the performances of each ML method did not change between training and testing datasets. Regarding CNN, the identified site-specific yield response functions were perfectly linear as we confirmed in the test results.

```{r table-optN-train, tab.cap = "Table D.1: Mean RMSE of EONR (kg/ha) estimation and profit-deficit ($/ha) by ML method and modeling scenario on training datasets"}
report_table_optN_train
```

Table D.2 shows the mean RMSE of yield prediction over the one thousand simulations using the training datasets as testing datasets. Interestingly, the performances of the prediction-oriented ML methods in predicting yields were relatively improved compared with the test results. Despite the improvement of prediction-oriented ML methods in yield prediction accuracy, it did not result in the more accurate EONR estimates. This also supports our argument that the model’s yield prediction accuracy has little to do with its EONR prediction accuracy.

```{r table-y-train, tab.cap = "Table D.2: Mean RMSE of yield prediction (kg/ha) by ML method and modeling scenario on training datasets"}
report_table_y_train
```
